---
title: "FARS Vignette"
author: "Axel Budde"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo = FALSE, include = FALSE}
library(fars)
library(dplyr)
library(maps)
```

## Purpose
This vignette provides an introduction to the fars package of functions for the analysis of data from the Fatality Analysis Reporting System (FARS).

## Installation
To install the fars package from GitHub:

```{r, eval = FALSE}
install.packages("devtools")
devtools::install_github("axelbudde/fars")
```

## Functions

The fars package offers three functions:

- `fars_read()`
- `fars_summarize_years()`
- `fars_map_state()`

## Examples

Examples are provided for each function, using FARS data from the years 2013 to 2015. The corresponding data files can be found in the `\extdata` folder.

## Reading FARS data

With `fars_read()` a file path can be defined. The function read the corresponding data file and returns a tibble.

```{r fars_read_example}
filename <- system.file("extdata/accident_2014.csv.bz2", package = "fars")
fars_read(filename)
```

## Summarizing data

With `fars_summarize_years()`summarizes the number of accidents by year and month by means of a pivot table. The function accepts a year or a range of years as arguments (integer).

```{r fars_summarize_years_example}
setwd(system.file("extdata", package = "fars"))
fars_summarize_years(2013:2015)
```

## Plotting accident locations

The function `fars_map_state()` plots accident locations in a certain state and year, using state index and year as arguments.

```{r fars_map_state_example}
setwd(system.file("extdata", package = "fars"))
fars_map_state(33, 2015)
```
